\textbf{Predicting virus-receptor mutant binding by molecular dynamics simulation}

### 2013-06-19 paper.tex
Claus's requested changes.

\subsection*{Steered molecular dynamics simulations}
To ensure a natural solution structure, we performed an initial equilibration phase to reduce the intrinsic crystallization distortions. First, we generated a physically realistic simulation world around the protein complex. The environment included an explicitly modeled bath of water, and a physiologically relevant concentration of ions. Next, with a fixed $\alpha$-carbon backbone we carried out successive rounds of energy minimization, followed by heating, and backbone release. Last,  with minimal restraints a short round of all-atom molecular dynamics \citep{Cuendet2008} was performed. We used the final state from each equilibrated system to restart another MD simulation. For SMD, at the centroid of several atoms a force was applied to a single point in GP1. By using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}. A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}. The force curves for each mutant were smoothed either by averaging over many replicates as in Figure~\ref{fig:force_curve} or by selecting an interpolation window for averaging. The latter approach was used for statistical analysis. In addition, to minimize the computational time in unproductive molecular dynamics, we shortened simulation time such that the interaction just barely dropped to zero near the end of each run. As seen in Figure~\ref{fig:force_curve}, this distance was relatively consistent among mutants. \textbf{Supplement 1 ($\leftarrow$ what is this? figure? video? text? table?)} illustrates the separation distance between peptide domains. The quantities MAF and AUC were derived from the force versus distances curves. Their summary statistics are reported in Table~\ref{tab:summary_statistics}. As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}.

@@

Although our computational predictions largely agreed with the available literature, some mutations that are reported to cause reduced infectivity in experimental systems (most notably vArg111Ala and Asn348Lys, discussed above) were indistinguishable from the WT complex in our simulations. We would like to emphasize here that we cannot expect perfect agreement. First, while we have shown that our method can distinguish individual point mutations, we do not know the limit of detection with our method. It is possible that some mutants display measurable phenotypic effects in experiments yet appear identical under our method. More extensive sampling or refinement of the simulation protocol could help to differentiate such mutants (see also next paragraph). Second, the available experimental data for the GP1/hTfR1 system were generally obtained from infection assays, not binding assays. There are several reasons other than binding due to which a viral mutation may cause a phenotypic difference in infectivity. \textbf{Shorten the following (till end of paragraph), make less critical to experimentalists. Just list the things other than binding that may matter.} First, across disconnected experiments there could be a gross difference in the GP1 concentration. Although western blots are a popular means of protein concentration normalization, when used in a quantitative manner they are notoriously difficult and unreliable. In addition, though western blots could show gross differences in protein concentration, viral binding may be very sensitive to small differences in fusion protein concentration.  Second, even if we trust protein expression normalization via western blot, it is not clear how much of the protein is grossly or partially misfolded. An additional analytical step with circular dichroism or an analogous technique may further clarify large-scale folding differences. Third, since we start with a bound structure, any changes that may dramatically affect the rate of association (different folds, trafficking issues, etc.) would be underestimated by our method.

>>>>

\subsection*{Steered molecular dynamics simulations}
To ensure a natural solution structure, we performed an initial equilibration phase to reduce the intrinsic crystallization distortions. First, we generated a physically realistic simulation world around the protein complex. The environment included an explicitly modeled bath of water, and a physiologically relevant concentration of ions. Next, with a fixed $\alpha$-carbon backbone we carried out successive rounds of energy minimization, followed by heating, and backbone release. Last,  with minimal restraints a short round of all-atom molecular dynamics \citep{Cuendet2008} was performed. We used the final state from each equilibrated system to restart another MD simulation. For SMD, at the centroid of several atoms a force was applied to a single point in GP1. By using a known spring constant, and pulling with a constant velocity, we could calculate the applied force during protein--protein dissociation \citep{Cuendet2008,Cuendet2011}. A typical averaged force curve comparison can be seen in Figure~\ref{fig:force_curve}. The force curves for each mutant were smoothed either by averaging over many replicates as in Figure~\ref{fig:force_curve} or by selecting an interpolation window for averaging. The latter approach was used for statistical analysis. In addition, to minimize the computational time in unproductive molecular dynamics, we shortened simulation time such that the interaction just barely dropped to zero near the end of each run. As seen in Figure~\ref{fig:force_curve}, this distance was relatively consistent among mutants. The supplementary movie visually illustrates the separation distance between peptide domains. The quantities MAF and AUC were derived from the force versus distances curves. Their summary statistics are reported in Table~\ref{tab:summary_statistics}. As we are more interested in the phenotypic impact of interface mutations we avoided many of the more physically rigorous, but technically complicated calculations that are possible with SMD \citep{Is2001A,Is2001B}.

@@

Although our computational predictions largely agreed with the available literature, some mutations that are reported to cause reduced infectivity in experimental systems (most notably vArg111Ala and Asn348Lys, discussed above) were indistinguishable from the WT complex in our simulations. We would like to emphasize here that we cannot expect perfect agreement. While we have shown that our method can distinguish individual point mutations, we do not know the limit of detection with our method. It is possible that some mutants display measurable phenotypic effects in experiments yet appear identical under our method. More extensive sampling or refinement of the simulation protocol could help to differentiate such mutants (see also next paragraph). In addition, the available experimental data for the GP1/hTfR1 system were generally obtained from infection assays, not binding assays. There are several reasons other than binding due to which a viral mutation may cause a phenotypic difference in infectivity. First, from infectivity experiments it is not clear the extent to which a protein is grossly or partially misfolded. An additional analytical step with circular dichroism or an analogous technique may further clarify large-scale folding differences. Second, since we start with a bound structure, any changes that may dramatically affect the rate of association (different folds, trafficking issues, etc.) would be underestimated by our method. Third, although it is tempting to think diminished infectivity should correlate strongly with decreased binding affinity, it is likely that dramatically increased binding affinity would prohibit cellular membrane detachment after viral invasion. As a result, the virus could not get to its replication compartment (which is the cytosol in the case of Machupo) and would not reproduce.
